---
title: "Class Presentation"
author: "Elizabeth Hawkey"
date: "11/5/2017"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1a. Data Wrangling
```{r}
library(lme4)
library(tidyverse)
library(broom)
library(dplyr)
library(psych)
library(tidyr)
library(merTools)
library(lavaan)
library(semTools)
library(semPlot)

PDS_stats1 <- read.csv(file =  "~/ejhawkey/STATS_resting_state_BRIEF.csv")
PDS_stats2 <- read.csv(file =  "~/ejhawkey/STATS_resting_state_BRIEF2.csv")

# often, stringsAsFactors = FALSE is useful
#names(PDS_stats2)

#Combine variables (at T3 there were two versions of the BRIEF: preschool and childhood)
#PDS_stats$T3gecrs_combined <- PDS_stats$T3gecrs + PDS_stats$T3gecrsP (tried this first it didn't work)

T3_gecrscombined <- PDS_stats2$T3_gecrs;   # start new column with one of the columns
inds <- which(is.na(PDS_stats2$T3_gecrsP) == FALSE);  # find non-NA entries in T3gecrsP
if (length(which(!is.na(T3_gecrscombined[inds]))) > 0) { stop("overlapping values??"); }
T3_gecrscombined[inds] <- PDS_stats2$T3_gecrsP[inds];  # stick the values in
PDS_stats2 <- data.frame(PDS_stats2, T3_gecrscombined);  # add new column in to existing dataframe
PDS_stats2$T3_gecrs <- PDS_stats2$T3_gecrscombined
describe(PDS_stats2$T3_gecrs)

#PDS_stats2$T1_agemonthsc <- scale(PDS_stats2$T1_agemonths, center = T, scale = F)

#Move data into a long format and a wide format.
wide_to_long_stats <- PDS_stats1 %>%
  gather(T1_sex:S3_DMNGEK6to10, key = "time", value = "value") 

wide_to_long_stats_1 <- wide_to_long_stats %>%
  separate(time, into = c("timepoint", "var"), sep = "_", convert = TRUE) %>%
  spread(key = var, value = value) %>%
  arrange(Subid) 
#View(wide_to_long_stats_2)

#Discuss ways to deal with layered longitudinal designs
#create a wave number to account for people starting at different waves
#this 
#dataframe <- dataframe %>%
  #group_by(Subid) %>%
  #mutate(wave = seq_len(n()))

#boxplot
library(ggplot2)
gg1 <-ggplot(data = wide_to_long_stats_1, aes(x = timepoint, y = gecrs)) +
  geom_boxplot(notch=FALSE, alpha=.8, outlier.colour = "gray", weight = 2, linetype = "longdash", colour="black") + 
  labs(x = "Timepoint", y = "EF Score") +
  xlim("T3", "T12", "T14") +
  ggtitle("Executive Function") +
  theme_minimal()
gg1


##Plot EF Composite score across 3 timepoints
gg2 <- ggplot(data = wide_to_long_stats_1,
          aes(x = timepoint, y = gecrs, group = Subid)) + geom_line(alpha = .5) +
          aes(colour = factor(Subid)) + guides(colour=FALSE) +
          xlim("T3", "T12", "T14") + 
          theme_minimal()
gg2


```


##More data wrangling. 
```{r}

#Move data with new combined variable into a long format again.
wide_to_long_stats <- PDS_stats2 %>%
  gather(T1_sex:T3_gecrscombined, key = "time", value = "value") 

wide_to_long_stats_2 <- wide_to_long_stats %>%
  separate(time, into = c("timepoint", "var"), sep = "_", convert = TRUE) %>%
  spread(key = var, value = value) %>%
  arrange(Subid) 
#View(wide_to_long_stats_2)


##Plot EF Composite score across 3 timepoints with combined T3 variable
gg3 <- ggplot(data = wide_to_long_stats_2,
          aes(x = timepoint, y = gecrs, group = Subid)) + geom_line(alpha = .5) +
          aes(colour = factor(Subid)) + guides(colour=FALSE) +
          xlim("T3", "T12", "T14") +
          theme_minimal()
gg3

#create a new variable to adjust EF to a scale that is closer to GE (to help models converge)
#wide_to_long_stats_2$gecrsconv <- as.numeric(wide_to_long_stats_2$gecrs/100)


```


#1. Run a series of MLM models using a categorical time variable. a) where the covariate only predicts the intercept b) predicts both intercept and slope c) is rescaled eg centering.
```{r}
##CATEGORICAL TIME VARIABLE
#a) only predicts the intercept
mod.1 <- lmer(gecrs ~ timepoint + (1|Subid), data = wide_to_long_stats_2)
summary(mod.1)

#b) predicts both intercept and slope 
mod.2 <- lmer(gecrs ~ timepoint + (timepoint|Subid), data = wide_to_long_stats_2)
summary(mod.2)

anova(mod.1, mod.2)

#WHAT IS THE PROBLEM HERE? UNEVEN GROUPS, CATEGORICAL PREDICTOR - CHARACTER?
describe.by(wide_to_long_stats_2$gecrs, group = wide_to_long_stats_2$timepoint)

##CONTINUOUS TIME VARIABLE
#adding time predicting intercept - fixed slope
mod.3 <- lmer(gecrs ~ agemonths + (1|Subid), data = wide_to_long_stats_2)
summary(mod.3)

#adding time - random slope
mod.4 <- lmer(gecrs ~ agemonths + (agemonths_c|Subid), data = wide_to_long_stats_2)
summary(mod.4)

anova(mod.3, mod.4)

```

#2a. Second order growth models - on BRIEF (using BRIEF composite scores (GEC = BRI + MI)
#Begin with a simple CFA to determine if latent variable is appropriate

For longitudinal models, occasion specific variance can lead to biased estimates. We want to separate the time specific variance from the overall construct variance. Or, we want to make sure that the time specific variance doesn’t make it appear that a construct is changing when really it is not.

```{r, warning = FALSE}

PDS_stats1 <- read.csv(file =  "~/ejhawkey/STATS_resting_state_BRIEF.csv")

GEC.model <- 'BRI_T3 =~ T3inhibrs + T3shftrs + T3emcnrs 
              MI_T3 =~ T3initirs + T3wrkmrs + T3plorgrs + T
fit.GEC <- cfa(GEC.model, data = PDS_stats, missing = "ML") 
summary(fit.GEC, fit.measures = TRUE) \
```

#2b. Build second order growth models 
```{r}
sec.order <- '
###define latent variables
EFT3 =~ NA*T3brirs + L1*T3brirs + L2*T3mirs 
EFT12 =~ NA*T12brirs + L1*T12brirs + L2*T12mirs 
EFT14=~ NA*T14brirs + L1*T14brirs + L2*T14mirs

### intercepts
T3brirs ~ t1*1
T3mirs ~ t2*1

T12brirs ~ t1*1
T12mirs ~ t2*1

T14brirs ~ t1*1
T14mirs ~ t2*1

#this is setting the means equal across waves

## correlated residuals across time
T3brirs ~~ T3mirs
T12brirs ~~ T12mirs
T14brirs ~~ T14mirs

## latent variable intercepts
EFT3 ~ 0*1
EFT12 ~ 0*1
EFT14 ~ 0*1

#model constraints for effect coding
## loadings must average to 1 (the three here changes to how many indicators you have; so change this based on the # of indicators)
L1 == 2 - L2
## means of indicators must average to 0 (in terms of the indicator means; )
t1 == 0 - t2

#the intercept and slope done with effect coding will give you the actual metric from your indicator variables (puts it in whatever metric you were measuring them in)

#final step is the normal growth model
i =~ 1*EFT3 + 1*EFT12 + 1*EFT14 
s =~ 0*EFT3 + 1*EFT12 + 2*EFT14 '

fit.sec.order <- growth(sec.order, data=PDS_stats)
summary(fit.sec.order, fit.measures=TRUE)

#Correlation between indicators for LV "EFT3"
m1= data.frame(PDS_stats$T3brirs, PDS_stats$T3mirs)
cor(m1, use="complete.obs")


#josh's code
#fit.sec.order <- growth(sec.order, data=long, missing = "ML")
#summary(fit.sec.order, fit.measures=TRUE)
```

3. Semplots for each of the models
```{r, echo=FALSE, warning=FALSE}

```
