---
title: "ALDA_HW3_EH"
author: "Elizabeth Hawkey"
date: "9/24/2017"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(lme4)
library(tidyverse)
library(broom)
library(tidyr)
library(merTools)
PDS_data_2 <- read.csv(file =  "~/Documents/PDS_project/Longitudinal_project/datasets/PDS_long_final.csv")
PDS_data_2$AGEMOSCAN <- as.numeric(PDS_data_2$AGEMOSCAN)
PDS_data_2$ADHDsum <- as.numeric(PDS_data_2$ADHDsum)
PDS_data_2$ADHD_INsum <- as.numeric(PDS_data_2$ADHD_INsum)
PDS_data_2$ADHD_HYIMsum<- as.numeric(PDS_data_2$ADHD_HYIMsum)
```

#1. Run a series of models using a time-invariant nominal covariate. a) where the covariate only predicts the intercept b) predicts both intercept and slope c) is rescaled eg centering. For all models, how does your model change from model to model. What is your final model?

```{r}
#nominal covariate = sex
#a) where the covariate only predicts the intercept

mod.1 <- lmer(ADHDsum ~ agemo_converted + sex + (1|Subid_fMRI), data=PDS_data_2)
summary(mod.1)

#b) predicts both intercept and slope 
mod.2 <- lmer(ADHDsum ~ agemo_converted + sex + (agemo_converted|Subid_fMRI), data=PDS_data_2)
summary(mod.2)

#c) is rescaled eg centering
sex_c <- scale(PDS_data_2$sex, center = T, scale = F)
PDS_data_2$sex_c <- as.numeric(sex_c)
mod.3 <- lmer(ADHDsum ~ agemo_converted + sex_c + (agemo_converted|Subid_fMRI), data=PDS_data_2)
summary(mod.3)

```
##The model fit improves (explains more of the variance) when a time variable is added, showing that sex is a more useful predictor of ADHD symptoms when the slopes are allowed to vary by subject. Centering sex does not improve the model.


#2. Introduce a time-invariant continuous covariate and run models a-c from #1.

```{r}
#continuous covariate = SES (Income_to_Need_Scan)
#a) where the covariate only predicts the intercept

mod.4 <- lmer(ADHDsum ~ Income_to_Need_Scan + (1|Subid_fMRI), data=PDS_data_2)
summary(mod.4)

#b) predicts both intercept and slope 
mod.5 <- lmer(ADHDsum ~ agemo_converted + Income_to_Need_Scan + (agemo_converted|Subid_fMRI), data=PDS_data_2)
summary(mod.5)

#c) is rescaled eg centering
Income_to_Need_Scan_c <- scale(PDS_data_2$ Income_to_Need_Scan , center = T, scale = F)
PDS_data_2$ Income_to_Need_Scan_c <- as.numeric(Income_to_Need_Scan_c)
mod.6 <- lmer(ADHDsum ~ agemo_converted + Income_to_Need_Scan_c + (agemo_converted|Subid_fMRI), data=PDS_data_2)
summary(mod.6)


```

#3. Graph both of your final models for the continuous and nominal models above.
Calculate confidence intervals around your estimates for your final models

```{r, echo=FALSE, warning=FALSE}
#nominal model graph


invisible(Predict.1 <- predictInterval(merMod = mod.3, newdata = PDS_data_2, level = 0.9, n.sims = 1000, 
                           stat = "median", include.resid.var = TRUE))

Predict.1$agemo_converted<- PDS_data_2$agemo_converted
Predict.1$Subid_fMRI <- PDS_data_2$Subid_fMRI

invisible(p.1 <- ggplot(aes(x = agemo_converted, y = fit), data = Predict.1) +
  geom_point(color = "cornflowerblue", size = .5) +
  labs(x="Age in Months", y="ADHD symptoms") + theme_bw() + 
  geom_line(data = Predict.1, aes(x = agemo_converted, y = fit, group=Subid_fMRI, color = Subid_fMRI), alpha = .4, size = .8) +
  stat_smooth(method = lm, se = T, size = 2, color = "cornflowerblue", alpha = .8) +
  ggtitle("Change in ADHD Symptoms Across Time") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_gray())
p.1 + theme(legend.position="none")

#continuous model graph
invisible(Predict.2 <- predictInterval(merMod = mod.6, newdata = PDS_data_2, level = 0.9, n.sims = 1000, 
                           stat = "median", include.resid.var = TRUE))

Predict.2$agemo_converted<- PDS_data_2$agemo_converted
Predict.2$Subid_fMRI <- PDS_data_2$Subid_fMRI

invisible(p.2 <- ggplot(aes(x = agemo_converted, y = fit), data = Predict.2) +
  geom_point(color = "blue", size = .5) +
  labs(x="Age in Months", y="ADHD Symptoms") + theme_bw() + 
  geom_line(data = Predict.2, aes(x = agemo_converted, y = fit, group=Subid_fMRI, color = Subid_fMRI), alpha = .4, size = .8) +
  stat_smooth(method = lm, se = T, size = 2, color = "blue", alpha = .8) +
  ggtitle("Change in ADHD Symptoms Across Time") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_gray())
p.2 + theme(legend.position="none")

###Calculate CI for final models
confint(mod.3, level = 0.95)
confint(mod.6, level = 0.95)
```

#4. Include both types of covariates in a single model. How does your interpretation of parameters change?

```{r}
mod.7 <- lmer(ADHDsum ~ agemo_converted + Income_to_Need_Scan_c + sex_c + (agemo_converted|Subid_fMRI), data=PDS_data_2)
summary(mod.7)

anova(mod.6, mod.7)
```
##Model 7 (including both covariates) appears to be a better fit since the AIC values are smaller.


#5. If you have one available, introduce a time-varying covariate.

```{r}
#time-varying covariate = MDD symptoms at each scan
mod.8 <- lmer(ADHDsum ~ agemo_converted + MDDCorescan + Income_to_Need_Scan_c + sex_c + (agemo_converted|Subid_fMRI), data=PDS_data_2)
summary(mod.8)

anova(mod.7, mod.8)

```

#6. Create a density plot of the random effects from your final model.

```{r}
library(merTools)
re.sim <- REsim(mod.7)
head(re.sim)
p.3 <- re.sim %>% 
  filter(term == "(Intercept)") 


ggplot(p.3, aes(mean)) +
  geom_density()
```

