---
title: "ALDA_HW2_EH"
author: "Elizabeth Hawkey"
date: "9/14/2017"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{}
library(lme4)
PDS_data_mlm <- read.csv(file = "~/Documents/PDS_project/Longitudinal_project/datasets/PDS_long_09_9_17_final.csv")
View(PDS_data)
```


#1.Run linear models on all of your subjects (a basic regression). What is the average intercept, the average slope?

```{load data}

```

#2. Now run a mlm/lmer model with only a random intercept. What is the ICC? What does residual variance look like compared to linear model? Create a graph to show this effect.

#3. Introduce a fixed slope term. What is the difference in terms of the fixed effects estimates between this estimate and the previous? Of the residual standard error? Create a graph to show both fixed effects estimates and the CIs around them.

#4. Run an additional model with a random slope. How does this change compare to the previous model? Should you keep the random slope or not?

#5. Interpret the correlation between the slope and the intercept.

#6. Create a density plot of the random effects from your final model.

#7. Create a catepilar plot of the random effects. Is there any person that seems odd in terms of a large standard errors around intercept and slope estimates?

#8. Create a plot of the trajectory, along with a spaghetti plot of each person’s individual slope. Set the alpha level (transparency) on the individual slopes to make them easier to see.

#9. Create a plot of the trajectory, along with a spagehtti plot of each person’s individual slope. Set the alpha level (transperancy) on the individual slopes to make them easier to see.



## Including Plots: You can also embed plots, for example:
```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
