---
title: "ALDA_HW3_EH"
author: "Elizabeth Hawkey"
date: "10/30/2017"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(lme4)
library(tidyverse)
library(broom)
library(tidyr)
library(merTools)
library(semTools)
library(semPlot)
library(dplyr)
library(psych)

PDS_stats <- read.csv(file =  "~/ejhawkey/STATS_resting_state_BRIEF3.csv")

```

#1. 1. Fit a measurement model to your constructs at one time point. Try out the different types of scaling discussed in class. What changes what stays the same?

##The fit statistics are the same across the two models but the covariances change.
```{r, warning = FALSE}

#Global executive compposit at T3
GEC.model <- 'BRI_T3 =~ inhibrs_3 + shftrs_3 + emcnrs_3 
              MI_T3 =~ initirs_3 + wrkmrs_3 + plorgrs_3 + orgmars_3 + monirs_3'


#Options for scaling (Largely irrelevant as to what scale is chosen. Serves to establish a point of reference in interpret other parameters):
#Marker variable (R default) Here you fix one factor loading to 1. All other loadings are relative to this loading.
fit.GEC.marker <- cfa(GEC.model, data = PDS_stats, missing = "ML")
summary(fit.GEC.marker, fit.measures = TRUE) 

semPaths(fit.GEC.marker, whatLabels = "est")


#Fixed variable: Here you fix the variance of the latent variable to 1 (standardized)
fit.GEC.fixed <- cfa(GEC.model, data = PDS_stats, std.lv = T, missing = "ML")
summary(fit.GEC.fixed, fit.measures = TRUE, standardized=TRUE) 

semPaths(fit.GEC.fixed, whatLabels = "std")

#Other option for scaling
#3. Effect coding. Here you constrain loading to average to 1. This will be helpful for us as we can then put the scale of measurement into our original metric. For longitudinal models this is helpful in terms of how to interpret the amount of change.



```

#2. What do the fit statistics say about your latent variable? Good/bad? Is your latent variable Just identified/saturdated, under identified or over identified?

##RMSEA = 0.144, SRMR = 0.043; Is this contradictory since one is >.10 and one is >0.08?
##CFI and TLI >.90, so this suggests that it is a good fit. 
## NO negative variances
##df = 19 In this model the knowns are greater than the unknowns (over identified)


#3. Fit a longitudinal CFA model where you:
#a) first correlate your latent factors across time and then 
##My would not converge

#b) a second model that predicts later times by a prevous time (ie auto regressive; t1 -> t2 -> t3). What are your conclusions? How does one differ from the other?

```{r, echo=FALSE}
# a) correlate latent factors across time
Long.GEC.model <- '
BRI_T3 =~ inhibrs_3 + shftrs_3 + emcnrs_3 
MI_T3 =~ initirs_3 + wrkmrs_3 + plorgrs_3 + orgmars_3 + monirs_3

BRI_T12 =~ inhibrs_12 + shftrs_12 + emcnrs_12 
MI_T12 =~ initirs_12 + wrkmrs_12 + plorgrs_12 + orgmars_12 + monirs_12

BRI_T14 =~ inhibrs_14 + shftrs_14 + emcnrs_14 
MI_T14 =~ initirs_14 + wrkmrs_14 + plorgrs_14 + orgmars_14 + monirs_14


## correlated residuals across time
inhibrs_3 ~~ inhibrs_12 + inhibrs_14
inhibrs_12 ~~ inhibrs_14
shftrs_3 ~~ shftrs_12 + shftrs_14
shftrs_12 ~~ shftrs_14
emcnrs_3 ~~ emcnrs_12 + emcnrs_14
emcnrs_12 ~~ emcnrs_14

initirs_3 ~~ initirs_12 + initirs_14
initirs_12 ~~ initirs_14
wrkmrs_3 ~~ wrkmrs_12 + wrkmrs_14
wrkmrs_12 ~~ wrkmrs_14
plorgrs_3 ~~ plorgrs_12 + plorgrs_14
plorgrs_12 ~~ plorgrs_14
orgmars_3 ~~ orgmars_12 + orgmars_14
orgmars_12 ~~ orgmars_14
monirs_3 ~~ monirs_12 + monirs_14
monirs_12 ~~ monirs_14'

fit.long.cfa <- cfa(Long.GEC.model, data=PDS_stats, std.lv=TRUE, missing = "ML")
summary(fit.long.cfa, standardized=TRUE, fit.measures=TRUE)
#semPaths(fit.long.cfa, whatLabels = "std")
#semPaths(fit.long.cfa, what = "std")

```

```{r}
#first look at correlations between variables
correlations <- select(PDS_stats, gecrscombined_3, gecrs_12, gecrs_14)
cor(correlations, use = "pairwise.complete.obs")
describe(correlations)
describe(correlations, na.rm = FALSE)

```

#4. 4. Fit a longitdinal growth model in SEM and in HLM. Compare and contrast the differences.

```{r}
PDS_stats_mlm <- read.csv(file =  "~/Documents/PDS_project/pds_dataset_info/wide_to_long_stats_2.csv")


##CATEGORICAL TIME VARIABLE
#a) only predicts the intercept
mod.1 <- lmer(gecrs ~ timepoint + (1|Subid), data = PDS_stats_mlm)
summary(mod.1)

#b) predicts both intercept and slope 
mod.2 <- lmer(gecrs ~ timepoint + (timepoint|Subid), data = PDS_stats_mlm)
summary(mod.2)
#this model does not converge either
#anova(mod.1, mod.2)

##CONTINUOUS TIME VARIABLE
#adding time predicting intercept - fixed slope
mod.3 <- lmer(gecrs ~ agemonths + (1|Subid), data = PDS_stats_mlm)
summary(mod.3)

#adding time - random slope
mod.4 <- lmer(gecrs ~ agemonths + (agemonths|Subid), data = PDS_stats_mlm)
summary(mod.4)


```
##longitudinal growth model isn't working

```{r}
# Global Executive Composite raw scores
# with intercept only 
 Intercept.only= ' i=~ 1*gecrscombined_3 + 1*gecrs_12 + 1*gecrs_14' 
 Intercept.only.fit= growth(Intercept.only, data = PDS_stats, missing= "ML")
 summary (Intercept.only.fit)
 semPaths(Intercept.only.fit, what = "paths", whatLabels= "est", layout = "tree")


# with a fixed slope
 fixed.slope= ' i=~ 1*T3gecrs_combined_conv + 1*T12gecrs_conv + 1*T14gecrs_conv
 s=~ 0*T3gecrs_combined_conv + 1*T12gecrs_conv + 2*T14gecrs_conv
 s ~~ 0*s'  #fixes slope
 fixed.slope.fit= growth(fixed.slope, data = PDS_stats, missing= "ML")
 inspect(fixed.slope.fit, "cov.lv")
 summary (fixed.slope.fit)
 semPaths(fixed.slope.fit, what = "paths", whatLabels= "est", layout = "tree")

# with a random slope

 random.intercept= ' i=~ 1*T3gecrs_combined_conv + 1*T12gecrs_conv + 1*T14gecrs_conv
 s=~ -1*T3gecrs_combined_conv + 0*T12gecrs_conv + 1*T14gecrs_conv'
 random.intercept.fit= growth(random.intercept, data = PDS_stats, missing= "ML")
 summary (random.intercept.fit)
 semPaths(random.intercept.fit, what = "paths", whatLabels= "est", layout = "tree")

# compare models 

 anova(Intercept.only.fit, fixed.slope)
 anova(fixed.slope, random.intercept.fit)
```


#5. Constrain the residual variances to be equal. Does this change the fit of your model?

```{r}


```

#6. Contrain your slope to be fixed, not random. How does this change your model?

```{r}

```

#7 Change the time metric in your SEM growth model. How does that change your estimates? Does it change your fit statistics?
```{r, echo=FALSE, warning=FALSE}

```

#8. Try a different type of estimation (see lavaan tutorial for details). How does that change your model?
```{r, echo=FALSE, warning=FALSE}

```


#9. Provide semplots for each of the models
```{r, echo=FALSE, warning=FALSE}

```

