---
title: "PDS_longitudinal_MLM"
author: "Elizabeth Hawkey"
date: "11/29/2017"
output:
  pdf_document: 
    latex_engine: xelatex
    #code_folding: "hide"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project Overview: Examine the time-varying relationship between brain network metric (global efficiency), and parent reported executive funciton scores across 3 waves of scan data.

###Step 1. Examine using a time-varying co-variate in lme4 using the Conners EF raw score and network gloal efficiency (GE) in FPN, DAN, SAL, CON. Examine multiple time variables and covariates

###Step 2. Compare GE to average within network connectivity. We would expect these to be very similar (no sig. difference)

###Step 3. Examine the time-varying relationship between EF and ADHD symptoms (we would expect this to be very significant), and MDD symptoms (possibly sig.)

##Data Management
```{r cars}
#load lme4 package: lme4 provides functions for fitting and analyzing mixed models: linear (lmer), 
#generalized linear (glmer) and nonlinear (nlmer.)
library(lme4)
library(tidyverse)
library(broom)
library(tidyr)
library(merTools)
library(semTools)
library(semPlot)
library(ggplot2)

#compute average efficiency metrics in SPSS prior to brining in dataset
#load in newest dataset from Rebecca Tillman
setwd("~/ejhawkey")
PDS_data_mlm <- read.csv("~/Documents/PDS_project/Longitudinal_project/datasets/PDS_long_9_8_17_MLM.csv")
#View(PDS_data_mlm)
#print(names(PDS_data_mlm))


#create a new variable to adjust age and EF to a scale that is closer to GE
#agemo_converted <- PDS_data_mlm$AGEMOSCAN/100
#CONP_EF_converted<- PDS_data_mlm$CONP_EF_T/100
#add back to dataframe
#PDS_data_mlm$agemo_converted <- as.numeric(agemo_converted)
#PDS_data_mlm$CONP_EF_converted <- as.numeric(CONP_EF_converted)

##Center predictor variables
#remember: centering is not the same as standardizing; 
#centering creates a zero mean; use when you don't need the units to be the same; you are changing the values not the scale
#standardizing: converting to z-scores; units change, 1 unit is now 1 SD; so that you can compare the variables in the same unit

##mean center by subject (so write a for loop for each subject to mean center each )
BRIEFgects_scan_c <- scale(PDS_data_mlm$BRIEFgects_scan, center = T, scale = F)
CONP_EF_RAW_c <- scale(PDS_data_mlm$CONP_EF_RAW, center = T, scale = F)
CONT_EF_RAW_c <- scale(PDS_data_mlm$CONT_EF_RAW, center = T, scale = F)
CONP_EF_T_c <- scale(PDS_data_mlm$CONP_EF_T, center = T, scale = F)
AGEMOSCAN_c <- scale(PDS_data_mlm$AGEMOSCAN, center = T, scale = F)
FPN_GE_K1to5_c <- scale(PDS_data_mlm$FPN_GE_K1to5, center = T, scale = F)	
FPN_GE_K6to10_c <- scale(PDS_data_mlm$FPN_GE_K6to10, center = T, scale = F)	
SAL_GE_K1to5_c <- scale(PDS_data_mlm$SAL_GE_K1to5, center = T, scale = F)	
SAL_GE_K6to10_c <- scale(PDS_data_mlm$SAL_GE_K6to10, center = T, scale = F)
DMN_GE_K1to5_c <- scale(PDS_data_mlm$DMN_GE_K1to5, center = T, scale = F)
DMN_GE_K6to10_c <- scale(PDS_data_mlm$DMN_GE_K6to10, center = T, scale = F)
CON_GE_K1to5_c <- scale(PDS_data_mlm$CON_GE_K1to5, center = T, scale = F)
CON_GE_K6to10_c <- scale(PDS_data_mlm$CON_GE_K6to10, center = T, scale = F)

#add the centered variables into the data frame
PDS_data_mlm$BRIEFgects_scan_c <- as.numeric(BRIEFgects_scan_c)
PDS_data_mlm$CONP_EF_RAW_c <- as.numeric(CONP_EF_RAW_c)
PDS_data_mlm$CONP_EF_T_c <- as.numeric(CONP_EF_T_c)
PDS_data_mlm$CONT_EF_RAW_c <- as.numeric(CONT_EF_RAW_c)
PDS_data_mlm$AGEMOSCAN_c <- as.numeric(AGEMOSCAN_c)
PDS_data_mlm$FPN_GE_K1to5_c <- as.numeric(FPN_GE_K1to5_c)
PDS_data_mlm$FPN_GE_K6to10_c <- as.numeric(FPN_GE_K6to10_c)
PDS_data_mlm$SAL_GE_K1to5_c <- as.numeric(SAL_GE_K1to5_c)
PDS_data_mlm$SAL_GE_K6to10_c <- as.numeric(SAL_GE_K6to10_c)
PDS_data_mlm$DMN_GE_K1to5_c <- as.numeric(DMN_GE_K1to5_c)
PDS_data_mlm$DMN_GE_K6to10_c <- as.numeric(DMN_GE_K6to10_c)
PDS_data_mlm$CON_GE_K1to5_c <- as.numeric(CON_GE_K1to5_c)
PDS_data_mlm$CON_GE_K6to10_c <- as.numeric(CON_GE_K6to10_c)

###center by person
#aggregate per subject all IVs and DVs
#cell_agg = cell %>% group_by(ID) %>%
#  summarize(acc_mean = mean)

###subset for scan 2, 3
#PDS_data_mlm.2.3 <- subset(PDS_data_mlm, scan_wave == 2:3)
#PDS_data_mlm.3 <- subset(PDS_data_mlm, scan_wave == 3)


write.csv(PDS_data_mlm, file = "~/Documents/PDS_project/Longitudinal_project/datasets/PDS_long_final.csv")

PDS_data_mlm <- read.csv(file = "~/Documents/PDS_project/Longitudinal_project/datasets/PDS_long_09_9_17_final.csv")

##START FROM HERE FOR ANALYSES
```

###setting up the models from class on 9/7
###lmer(y ~ 1 + time + (1 + time | subjects), data = data)
###lmer(y ~ time + (time|Subid), data = data)  - both are equivalent
###outside the parentheses: fixed effects
###terms inside the parentheses (how we specify random effects 1 = not every indiv. will have the same intercept value
### pipe (|) denotes nesting variable (for us it is subject level identifier))


## Global Efficiency models with Conners EF prediciton GE
###Salience Network
```{r, warning = FALSE}
##SAL
mod.SAL.GE1a <- lmer(SAL_GE_K1to5 ~ agemo_converted + sex + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.SAL.GE1a)
mod.SAL.GE1b <- lmer(SAL_GE_K1to5 ~ agemo_converted*CONP_EF_RAW_c + sex + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.SAL.GE1b)

PDS_data_mlm.na.rm <- subset(PDS_data_mlm, SAL_GE_K1to5 > 0)
PDS_data_mlm.na <- subset(PDS_data_mlm.na.rm, CONP_EF_RAW_c >0)

##just main effects alone; very similar to main effects in interaction model; no need to report this
#mod.SAL.GE1c <- lmer(SAL_GE_K1to5 ~ CONP_EF_RAW_c + (agemo_converted | Subid_fMRI), data=PDS_data_mlm.na.rm)
#summary(mod.SAL.GE1c)

mod.SAL.GE2a <- lmer(SAL_GE_K6to10 ~ agemo_converted + sex + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.SAL.GE2a)
mod.SAL.GE2b <- lmer(SAL_GE_K6to10 ~ agemo_converted*CONP_EF_RAW_c + sex + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.SAL.GE2b)

```


###Cingulo-opercular Network
```{r, eval=FALSE}
##CON
mod.CON.GE1a <- lmer(CON_GE_K1to5 ~ agemo_converted+ sex + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.CON.GE1a)
mod.CON.GE1b <- lmer(CON_GE_K1to5 ~ agemo_converted*CONP_EF_RAW_c + sex + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.CON.GE1b)

mod.CON.GE2 <- lmer(CON_GE_K6to10 ~ agemo_converted + sex + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.CON.GE2)
mod.CON.GE2 <- lmer(CON_GE_K6to10 ~ agemo_converted*CONP_EF_RAW_c + sex + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.CON.GE2)
```


###Frontoparietal Network
```{r}
##FPN
mod.FPN.GE1a <- lmer(FPN_GE_K1to5 ~ agemo_converted + sex + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.FPN.GE1a)
mod.FPN.GE1b <- lmer(FPN_GE_K1to5 ~ agemo_converted*CONP_EF_RAW_c + sex + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.FPN.GE1b)

mod.FPN.GE2a <- lmer(FPN_GE_K6to10 ~ agemo_converted + sex + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.FPN.GE2a)
mod.FPN.GE2b <- lmer(FPN_GE_K6to10 ~ agemo_converted*CONP_EF_RAW_c + sex + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.FPN.GE2b)

#anova(mod.FPN.GE2a, mod.FPN.GE2b)
```


###Default Mode Network
```{r}
#DMN
mod.DMN.GE1a <- lmer(DMN_GE_K1to5 ~ agemo_converted + sex + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.DMN.GE1a)
mod.DMN.GE1b <- lmer(DMN_GE_K1to5 ~ agemo_converted*CONP_EF_RAW_c + sex + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.DMN.GE1b)

mod.DMN.GE2a <- lmer(DMN_GE_K6to10 ~ agemo_converted + sex + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.DMN.GE2a)
mod.DMN.GE2b <- lmer(DMN_GE_K6to10 ~ agemo_converted*CONP_EF_RAW_c + sex + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.DMN.GE2b)
```

###AVERAGE WITHIN CONNECTIVITY
```{r}
##AVERAGE WITHIN CONNECTIVITY
mod.SAL <- lmer(SAL_SAL ~ agemo_converted*CONP_EF_RAW_c + sex + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.SAL)

mod.CON <- lmer(CON_CON ~ agemo_converted*CONP_EF_RAW_c + sex + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.CON)

mod.DMN <- lmer(DMN_DMN ~ agemo_converted*CONP_EF_RAW_c + sex + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.DMN)

mod.FPN <- lmer(FPN_FPN ~ agemo_converted*CONP_EF_RAW_c + sex + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.FPN)
```


##ADHD 
```{r}
#ADHDsum; parent 
mod.ADHD1 <- lmer(ADHDsum ~ agemo_converted + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.ADHD1)
anova(mod.ADHD1)
mod.ADHD2 <- lmer(ADHDsum ~ agemo_converted*CONP_EF_RAW_c + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.ADHD2)
mod.ADHD3 <- lmer(ADHDsum ~ agemo_converted*CONP_EF_RAW_c + sex + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.ADHD3)

#ADHD_INsum
mod.ADHDIN <- lmer(ADHD_INsum ~ agemo_converted*CONP_EF_RAW_c + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.ADHDIN)

#ADHD_HYIMsum
mod.ADHDHY <- lmer(ADHD_HYIMsum ~ agemo_converted*CONP_EF_RAW_c + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.ADHDHY)

##EF and MDD
mod.MDD1 <- lmer(MDDCorescan ~ agemo_converted + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.MDD1)
mod.MDD2 <- lmer(MDDCorescan ~ agemo_converted*CONP_EF_RAW_c + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.MDD2)
mod.MDD3 <- lmer(MDDCorescan ~ agemo_converted*CONP_EF_RAW_c + sex + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.MDD3)

##Next do GE predicting ADHD
#ADHDsum

mod.ADHD.SAL1 <- lmer(ADHDsum ~ agemo_converted*SAL_GE_K1to5_c + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.ADHD.SAL1)
mod.ADHD.SAL2 <- lmer(ADHDsum ~ agemo_converted*SAL_GE_K6to10_c + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.ADHD.SAL2)

mod.ADHD.CON1 <- lmer(ADHDsum ~ agemo_converted*CON_GE_K1to5_c + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.ADHD.CON1)
mod.ADHD.CON2 <- lmer(ADHDsum ~ agemo_converted*CON_GE_K6to10_c + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.ADHD.CON2)

mod.ADHD.FPN1 <- lmer(ADHDsum ~ agemo_converted*FPN_GE_K1to5_c + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.ADHD.FPN1)
mod.ADHD.FPN2 <- lmer(ADHDsum ~ agemo_converted*FPN_GE_K6to10_c + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.ADHD.FPN2)

mod.ADHD.DMN1 <- lmer(ADHDsum ~ agemo_converted*DMN_GE_K1to5_c + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.ADHD.DMN1)
mod.ADHD.DMN2 <- lmer(ADHDsum ~ agemo_converted*DMN_GE_K6to10_c + (agemo_converted | Subid_fMRI), data=PDS_data_mlm)
summary(mod.ADHD.DMN2)
```


```{r}
##output sig. models together in one table; breatk it up by analysis type?
#library(sjPlot)
#sjt.lm(mod.ADHD.SAL1 , mod.ADHD.SAL2, file = "sjt.ADHD.SAL.doc")
#sjt.lm(adhdmodel1, adhdmodel2, adhdmodel3, adhdmodel4, adhdmodel5, file = "sjt_adhd.doc")

```


### Data Visualization
```{r pressure, echo=FALSE, warning=FALSE}
##make pretty plots
plot.connersEF <- ggplot(PDS_data_mlm, aes(x = AGEMOSCAN, y = CONP_EF_RAW_c, group = Subid_fMRI)) +
  geom_point() + 
  geom_line(alpha = .5) + 
  aes(colour = factor(Subid_fMRI)) + 
  guides(colour=FALSE) +
  theme_classic()
plot.connersEF

plot.adhdsum <- ggplot(PDS_data_mlm, aes(x = AGEMOSCAN, y = ADHDsum, group = Subid_fMRI)) +
  geom_point() + 
  geom_line(alpha = .5) + 
  aes(colour = factor(Subid_fMRI)) + 
  guides(colour=FALSE) +
  theme_classic()
plot.adhdsum
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

###Additional analyses: Teacher
```{r}
#ADHDsum; teacher
#PDS_data_mlm_T_na <- subset(PDS_data_mlm, CONT_EF_RAW >= 0)

#mod.T.ADHD0 <- lm(ADHDsum ~ CONT_EF_RAW + sex, data = PDS_data_mlm_teacher)
#summary(mod.T.ADHD0)
#mod.T.ADHD1 <- lmer(ADHDsum ~ agemo_converted + (1 | Subid_fMRI), data=PDS_data_mlm_T_na)
#summary(mod.T.ADHD1)
#mod.T.ADHD2 <- lmer(ADHDsum ~ CONT_EF_RAW + (agemo_converted | Subid_fMRI), data = PDS_data_mlm_T_na)
#summary(mod.T.ADHD2)
#mod.T.ADHD3 <- lmer(ADHDsum ~ agemo_converted*CONT_EF_RAW + sex + (agemo_converted | Subid_fMRI), data=PDS_data_mlm_T_na)
#summary(mod.T.ADHD3)
#this isn't working
```

