---
title: "Conditional_Models_EH"
author: "Elizabeth Hawkey"
date: "10/5/2017"
output: 
  pdf_document: 
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, results = 'hide', message = F, warning = F}
library(psych)
library(sjPlot)
library(broom)
library(lme4)
library(MuMIn)
library(merTools)
library(reghelper)
library(stargazer)
library(lsmeans)
library(multcompView)
library(plyr)
library(tidyverse)

PDS_data_2 <- read.csv(file =  "~/Documents/PDS_project/Longitudinal_project/datasets/PDS_long_final.csv")
PDS_data_2$AGEMOSCAN <- as.numeric(PDS_data_2$AGEMOSCAN)
PDS_data_2$ADHDsum <- as.numeric(PDS_data_2$ADHDsum)
PDS_data_2$ADHD_INsum <- as.numeric(PDS_data_2$ADHD_INsum)
PDS_data_2$ADHD_HYIMsum<- as.numeric(PDS_data_2$ADHD_HYIMsum)

mod.8 <- lmer(ADHDsum ~ agemo_converted + MDDCorescan + Income_to_Need_Scan_c + sex_c + (agemo_converted|Subid_fMRI), data=PDS_data_2)
summary(mod.8)
```

#Below is code that *should* work for all models. Just run the function and save it as an `R` object. You can use this with `papaja` and the `apa_table()` function pretty easily. The trick is that if you are not using the papaja template, the proper LaTeX packages may not be loaded. You can get around this by attaching a .tex file calling the packages under "in_header: header.tex" in your YAML header. The YAML header of this .Rmd file contains the necessary syntax and the header.tex file with the proper packages.
 
```{r, echo = T, results = 'hide'}
## here's some code to make a table. You shouldn't need to modify anything here 
# unless you add additional random effects terms
## fixed effects first ##
table_fun <- function(model){
  fixed <- broom::tidy(model) %>% filter(group == "fixed") %>%
    select(term, estimate) 
  ## add random effects ##
  rand <- broom::tidy(model) %>% filter(group != "fixed") %>%
    select(term, estimate)
  ## get confidence intervals ##
  CI <- data.frame(confint.merMod(model, method = "boot", nsim = 10)) %>%
    mutate(term = rownames(.)) %>% setNames(c("lower", "upper", "term"))
  
  ## Get ICC & R2 values ##
  ICC <- reghelper::ICC(model)
  R2 <- MuMIn::r.squaredGLMM(model)
  
  ## format the fixed effects
  fixed <- fixed %>% left_join(CI %>% filter(!grepl(".sig", term))) %>%
    mutate(type = "Fixed Parts")
  
  rand <- rand %>%
    mutate(estimate = ifelse(grepl("cor", term) == T, estimate, estimate^2),
          term = mapvalues(term, unique(term), 
           c("$\\tau_{00}$", "$\\tau_{11}$", "$\\tau_{10}$", "$\\hat{\\sigma^2}$"))) %>%
    left_join(
      CI %>% filter(grepl(".sig", term)) %>%
        mutate(term = mapvalues(term, unique(term), 
            c("$\\tau_{00}$", "$\\tau_{10}$", "$\\tau_{11}$", "$\\hat{\\sigma^2}$")),
            lower = lower^2, upper = upper^2)) %>%
    mutate(type = "Random Parts")
  
  mod_terms <- tribble(
    ~term, ~estimate, ~type,
    "ICC", ICC, "Model Terms",
    "$R^2_m$", R2[1], "Model Terms",
    "$R^2_c$", R2[2], "Model Terms"
  )
  
  tab <- fixed %>%
    full_join(rand) %>%
    mutate(CI = sprintf("(%.2f, %.2f)", lower, upper)) %>%
    select(-lower, -upper) %>%
    full_join(mod_terms) %>%
    mutate(estimate = sprintf("%.2f", estimate)) %>%
    dplyr::rename(b = estimate) %>%
    select(type, everything())
  return(tab)
}
# you can use this with papaja and the apa_table function pretty easily
# the trick is that if you are not using the papaja template, the proper
# LaTeX packages may not be loaded. You can get around this by attaching
# a .tex file calling the packages under "in_header: header.tex" in your YAML
# header the YAML header of this .Rmd file contains the necessary syntax and 
# the header.tex file with the proper packages

tab <- table_fun(mod.8)
```


```{r}
options(knitr.kable.NA = '')
knitr::kable(tab, caption = "Ugly MLM Table Example")
```

### Alternative: `papaja` + `apa_table()`  

```{r, results = 'asis'}
papaja::apa_table(tab %>% select(-type),caption = "papaja MLM Table Example", 
    na_string = "", stub_indents = list(Fixed = c(1:4), Random = c(5:8), Summary = c(9:11)),
    col_spanners = list(`Sensation Seeking` = c(2,3)))
```

```{r, echo=FALSE, warning=FALSE}
#continuous model graph
invisible(Predict.2 <- predictInterval(merMod = mod.8, newdata = PDS_data_2, level = 0.9, n.sims = 1000, 
                           stat = "median", include.resid.var = TRUE))

Predict.2$agemo_converted<- PDS_data_2$agemo_converted
Predict.2$Subid_fMRI <- PDS_data_2$Subid_fMRI

invisible(p.2 <- ggplot(aes(x = agemo_converted, y = fit), data = Predict.2) +
  geom_point(color = "blue", size = .5) +
  labs(x="Age in Months", y="ADHD Symptoms") + theme_bw() + 
  geom_line(data = Predict.2, aes(x = agemo_converted, y = fit, group=Subid_fMRI, color = Subid_fMRI), alpha = .4, size = .8) +
  stat_smooth(method = lm, se = T, size = 2, color = "blue", alpha = .8) +
  ggtitle("Change in ADHD Symptoms - Controlling for MDD Symptoms") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_gray())
p.2 + theme(legend.position="none")
```

